---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "ragg_png",
  warning = FALSE,
  error = FALSE,
  message = FALSE
)
```

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/vidonne/unhcrthemes/workflows/R-CMD-check/badge.svg)](https://github.com/vidonne/unhcrthemes/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/unhcrthemes)](https://CRAN.R-project.org/package=unhcrthemes)
<!-- badges: end -->

# unhcrthemes

The **unhcrthemes** package provides a `ggplot` theme and a set of colour palettes for making charts and graphics based on [UNHCR Data Visualization Guidelines](https://www.unhcr.org/brand/wp-content/uploads/sites/89/2021/11/UNHCR_Data_Visualization_Guidelines.pdf). The goal of this package is to ease and speed up the creation of charts, while promoting the UNHCR visual identity with a predefined `ggplot` theme, as well as a set of colour palettes and scales. 

## Installation

This package is not on yet on CRAN and to install it, you will need the remotes package.

```{r install, eval = FALSE}
install.packages("remotes")
remotes::install_github("vidonne/unhcrthemes")
```

## Content

A package with all necessary elements to quickly implement [UNHCR Brand style](https://www.unhcr.org/brand/how-we-look/) in your statistical products and data stories:

1. Adjusted `ggplot2` theme
2. A series of color palette for:
    - A **categorical palette** for UNHCR main data visualization colors
    - A **categorical palette** for people of concern to UNHCR categories
    - A **categorical palette** for geographical regional divisions of UNHCR
    - Six **sequential color palettes** for all the main data visualization colors
    - Two recommended **diverging color palette**

## Fonts

UNHCR uses **Lato** as its main font for publications and data visualizations. **Arial** can also be used as a fallback option if it is not possible to install **Lato** on your computer.

**Lato** needs to be installed on your computer and registered with R. Install [Lato](https://fonts.google.com/specimen/Lato), then run the following code to make it accessible in R.

**TO REVIEW**

```{r fonts}
# install.packages('showtext', dependencies = TRUE)
library(showtext)

# Check the current search path for fonts
allfontpath <- font_paths()

# syntax: font_add(family = "<family_name>", regular = "/path/to/font/file")
font_add("Lato", regular = "Lato-Regular.ttf",  bold = "Lato-Bold.ttf", italic = "Lato-Italic.ttf")

# automatically use showtexts for new devices
showtext::showtext_auto()
```

## Usage

```{r load}
library(ggplot2) # or library(tidyverse)
library(unhcrthemes)
```

### Base ggplot2 theme

```{r plot-theme}
ggplot(data = mtcars, mapping = aes(factor(cyl))) +
  geom_bar() + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Number of Cylinders",
       y = "Count", title = "Bar chart example",
       subtitle = "Simple plot to showcase theme_unhcr default style",
       caption = "Data from mtcars") +
  theme_unhcr(grid = "Y")
```

### UNHCR color palette

All recommended data visualization colors are accessible as **palettes** or **scales** (color/fill).

```{r palette}
display_unhcr_all()
```

### Base theme and color scale

```{r plot-theme-fill}
ggplot(data = mtcars, mapping = aes(factor(cyl))) +
  geom_bar(fill = unhcr_pal(n=1, name = "pal_unhcr")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Number of Cylinders",
       y = "Count", title = "Bar chart example with UNHCR colors",
       subtitle = "Simple plot to showcase theme_unhcr default style",
       caption = "Data from mtcars") +
  theme_unhcr(grid = "Y")
```

```{r plot-theme-color}
ggplot(mtcars, aes(mpg, wt)) +
  geom_point(aes(color=factor(gear))) +
  scale_color_unhcr_d(palette = "pal_unhcr") +
  labs(x = "Number of Cylinders",
       y = "Count", title = "Scatter plot example with UNHCR colors",
       subtitle = "Simple plot to showcase theme_unhcr default style",
       caption = "Data from mtcars") +
  theme_unhcr(axis = FALSE)

```

```{r child="man/rmdfragments/unhcrverse.Rmd"} 
```

## Meta

- Please [report any issues or bugs](https://github.com/vidonne/unhcrthemes/issues).
- License: MIT
- Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
