---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.retina = 2,
  dpi = 300,
  dev = "ragg_png"
)
```

# unhcrthemes <img src='man/figures/logo.svg' align="right" alt="unhcrthemes package hex sticker" width="140" />

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![R-CMD-check](https://github.com/unhcr-dataviz/unhcrthemes/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/unhcr-dataviz/unhcrthemes/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/unhcrthemes)](https://CRAN.R-project.org/package=unhcrthemes)
<!-- badges: end -->

Simplify the creation of your data visualizations with the `{unhcrthemes}` R package. It is designed to align your plots with [UNHCR's data visualization](https://dataviz.unhcr.org/general_guidance/) recommendations. This package offers a comprehensive `{ggplot2}` theme, including built-in customizations and thoughtfully crafted color palettes.

## Features

- A consistent and appealing `{ggplot2}` theme that matches [UNHCR Data Visualization Guidelines](https://dataviz.unhcr.org/general_guidance/).
- A collection of color palettes ensuring your plots are consistent with [UNHCR's color scheme](https://dataviz.unhcr.org/general_guidance/).
- Easy-to-use functions to apply the theme and color palettes to your `{ggplot2}` plots.

## Installation

Get started by installing the `{unhcrthemes}` package from CRAN with:

```{r install, eval=FALSE}
install.packages("unhcrthemes")
```

Or retrieve the development version from Github using the `{pak}` package:

```{r pak, eval=FALSE}
# If pak is not yet installed, uncomment the following line:
# install.packages("pak")

pak::pkg_install("unhcr-dataviz/unhcrthemes")
```

## Usage

```{r load, message = FALSE, warning = FALSE}
# Load required packages
library(ggplot2)
library(unhcrthemes)
```

### Base `{ggplot2}` theme

```{r base-plot, message = FALSE, warning = FALSE, out.width="90%", fig.align="center"}
# Apply the theme_unhcr() to your plot
ggplot(
  mtcars,
  aes(x = hp, y = mpg)
) +
  geom_point() +
  labs(
    title = "Simple scatterplot with ggplot",
    subtitle = "This plot shows the default style of theme_unhcr()",
    x = "Horsepower (hp)",
    y = "Fuel efficiency (mpg)",
    caption = "Source: mtcars"
  ) +
  theme_unhcr()
```

### Font

One of the two officially recommended fonts outlined in the [UNHCR Data Visualization Guidelines](https://dataviz.unhcr.org/general_guidance/) is [`Lato`](https://fonts.google.com/specimen/Lato). To ensure optimal functionality of the `{unhcrthemes}` package, please make sure that the `Lato` font is installed on your device prior to usage.

### Color palettes

Choose from a variety of carefully curated color palettes that harmonize with UNHCR's visual identity.

```{r palette-all, message = FALSE, warning = FALSE, out.width="90%", fig.align="center", fig.height=3.2}
# Display all unhcr color palette
display_unhcr_all()
```

```{r palette-blue, message = FALSE, warning = FALSE, out.width="90%", fig.align="center", fig.height=2}
# Or select a specific one and limit the number of colors
display_unhcr_pal(n = 3, "pal_blue")
```

### Scales (color/fill)

```{r scale-plot, message = FALSE, warning = FALSE, out.width="90%", fig.align="center"}
# Add color scale to your plot
ggplot(
  mtcars,
  aes(x = hp, y = mpg, color = as.factor(cyl))
) +
  geom_point(alpha = .8) +
  labs(
    title = "Simple scatterplot with ggplot",
    subtitle = "This plot shows the default style of theme_unhcr()",
    x = "Horsepower (hp)",
    y = "Fuel efficiency (mpg)",
    color = "Number of cylinder",
    caption = "Source: mtcars"
  ) +
  scale_color_unhcr_d() +
  theme_unhcr(legend_title = TRUE)
```

### Geoms (text/label)

```{r geom-plot, message = FALSE, warning = FALSE, out.width="90%", fig.align="center", fig.height=4}
# Default Lato font also apply for geom_text and label
dplyr::count(mpg, class) |>
  dplyr::arrange(n) |>
  dplyr::mutate(class = factor(class, levels = class)) |>
  ggplot(aes(n, class)) +
  geom_col(fill = unhcr_pal(n = 1, "pal_blue")) +
  geom_text(aes(label = n), hjust = 0, nudge_x = 1) +
  labs(
    title = "Simple bar chart with labels using ggplot",
    subtitle = "This plot shows the default style of theme_unhcr()",
    caption = "Source: mtcars"
  ) +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  theme_unhcr(grid = FALSE, axis = "y", axis_title = FALSE, axis_text = "y")
```

Explore more examples of `{unhcrthemes}` in action on the [package vignette](https://unhcr-dataviz.github.io/unhcrthemes/articles/intro-unhcrthemes.html) or [UNHCR Data Visualization Platform](https://dataviz.unhcr.org/tools/r/).

## Additional UNHCR Tools

- [`{refugees}`](https://github.com/PopulationStatistics/refugees): A comprehensive package for accessing [UNHCR Refugee Population Statistics Database](https://www.unhcr.org/refugee-statistics/download/).
- [`{unhcrdown}`](https://github.com/unhcr-dataviz/unhcrdown): UNHCR branded templates for [R Markdown](https://rmarkdown.rstudio.com/).
- [`quarto-revealjs-unhcr`](https://github.com/unhcr-dataviz/quarto-revealjs-unhcr): A [Quarto](https://quarto.org/) extension to create [UNHCR branded](https://www.unhcr.org/brand) `reveal.js` presentations.
- [`quarto-html-unhcr`](https://github.com/unhcr-dataviz/quarto-html-unhcr): A [Quarto](https://quarto.org/) extension to create [UNHCR branded](https://www.unhcr.org/brand) `HTML` document.
- [Data Visualization Platform](https://dataviz.unhcr.org/): A collection of recommendations on selecting the best type of chart, general dataviz design, tools and tutorials and a selection of data visualization products made by UNHCR.

## Acknowledgements

We extend our gratitude to the creators of the [`{hrbrthemes}`](https://github.com/hrbrmstr/hrbrthemes), [`{rcartocolor}`](https://github.com/Nowosad/rcartocolor) and [`{cowplot}`](https://github.com/wilkelab/cowplot/) packages that influenced the development of `{unhcrthemes}`.

## Contribution and Code of Conduct

Contributions to `{unhcrthemes}` are highly valued. To ensure a welcoming and inclusive community, we follow our [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). If you have suggestions, uncover bugs, or envision new features, kindly submit an [issue on GitHub](https://github.com/unhcr-dataviz/unhcrthemes/issues). To contribute code, don't hesitate to fork the repository and create a pull request.

## License

This package is distributed under the [MIT License](https://github.com/unhcr-dataviz/unhcrthemes/blob/master/LICENSE.md).
