---
title: "UNHCR branded charts with unhcrthemes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{UNHCR branded charts with unhcrthemes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(ragg)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  fig.showtext = TRUE,
  dpi = 128
)
```

## Packages

```{r setup}
library(unhcrthemes)
library(tidyverse)
library(unhcrthemes)
library(unhcrdatapackage)
library(scales)
library(lubridate)
library(directlabels)
library(ggforce)
library(ggtext)
library(treemapify)
```

## Column chart

```{r bar1, fig.width = 5.5, fig.height = 3}
df1 <- tribble(
~year, ~nidp,
"2010",24.98,
"2011",22.44,
"2012",26.39,
"2013",33.34,
"2014",37.87,
"2015",40.52,
"2016",40.29,
"2017",39.98,
"2018",41.34,
"2019",45.70,
"2020",48.03)


ggplot(df1) +
  geom_col(aes(x = year, y = nidp),
           fill = unhcr_pal(n = 1, "pal_blue"),
           width = 0.8) +
  labs(title = "Globalement IDP displacement | 2010 - 2020",
       subtitle = "Number of people (in million)",
       x = "",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_y_continuous(expand = expansion(c(0, 0.1))) +
  theme_unhcr(font_size = 9,
              font_family = "Lato",
              grid = "Y")
```

## Column chart with data label


```{r bar1_alt, fig.width = 5.5, fig.height = 3}
ggplot(df1) +
  geom_col(aes(x = year, y = nidp),
           fill = unhcr_pal(n = 1, "pal_blue"),
           width = 0.8) +
  geom_text(aes(x = year, y = nidp, label = nidp),
            vjust = -1,
            size = 3) +
  labs(title = "Globalement IDP displacement | 2010 - 2020",
       subtitle = "Number of people (million)",
       x = "",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_y_continuous(expand = expansion(c(0, 0.1))) +
  theme_unhcr(font_size = 9,
              font_family = "Lato",
              grid = FALSE,
              axis = "X") +
  theme(axis.text.y = element_blank())
```

## Bar chart

```{r col1, fig.width = 5.5, fig.height = 5}
df2 <- tribble(
  ~country,~displaced,
  "Eritrea",522000,
  "Central African Rep.",642000,
  "Sudan",788000,
  "Somalia",815000,
  "DRC",840000,
  "Myanmar",1103000,
  "South Sudan",2189000,
  "Afghanistan",2595000,
  "Venezuela",4027000,
  "Syria",6690000)

df2 |>
  ggplot() +
  geom_col(aes(x = displaced/1e6,
               y = reorder(country, displaced)),
           fill = unhcr_pal(n = 1, "pal_blue"),
           width = 0.8) +
  labs(title = "People displaced across borders by country of origin | 2020",
       x = "Number of people (million)",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_x_continuous(expand = expansion(c(0, 0.1)),
                     breaks = pretty_breaks()) +
  theme_unhcr(font_size = 9,
              font_family = "Lato",
              grid = "X",
              axis = "y")
```

## Bar chart with data label

```{r col1_alt, fig.width = 5.5, fig.height = 5}
df2 |>
  ggplot() +
  geom_col(aes(x = displaced/1e6,
               y = reorder(country, displaced)),
           fill = unhcr_pal(n = 1, "pal_blue"),
           width = 0.8) +
  geom_text(aes(x = displaced/1e6,
                y = reorder(country, displaced),
                label = round(displaced/1e6, 2)),
            hjust = -0.5,
            size = 3) +
  labs(title = "People displaced across borders by country of origin | 2020",
       subtitle = "Number of people (million)",
       x = "",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  theme_unhcr(font_size = 9,
              font_family = "Lato",
              grid = FALSE,
              axis = "Y") +
  theme(axis.text.x = element_blank())
```


## Grouped column chart

```{r bar2, fig.width = 5.5, fig.height = 3}
df3 <- tribble(
  ~year,~EHAGL,~RBSA,~RBWCA,
  "2017",7,4,3,
  "2018",9,5,4,
  "2019",8,6,5,
  "2020",10,6,6)


df3 |>
  pivot_longer(cols = -year) |>
  ggplot() +
  geom_col(aes(year, value, fill = name),
           position = position_dodge(width = 0.7),
           width = 0.6) +
  scale_fill_unhcr_d(palette = "pal_unhcr", nmax = 3, order = c(2, 3, 1)) +
  labs(title = "IDPs of concern to UNHCR in Africa region | 2017-2020",
       x = "",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_y_continuous(expand = expansion(c(0, 0.2)),
                     breaks = pretty_breaks()) +
  theme_unhcr(font_size = 9, grid = "Y") +
  theme(axis.title.y = element_blank())
```

## Grouped column chart with data label

```{r bar2_alt, fig.width = 5.5, fig.height = 3}
df3 |>
  pivot_longer(cols = -year) |>
  ggplot() +
  geom_col(aes(x = year, y = value, fill = name),
           position = position_dodge(width = 0.7),
           width = 0.6) +
  geom_text(aes(year, value, group = name, label = value),
            position = position_dodge(width = 0.7),
            vjust = -1,
            size = 3) +
  scale_fill_unhcr_d(palette = "pal_unhcr", nmax = 3, order = c(2, 3, 1)) +
  labs(title = "IDPs of concern to UNHCR in Africa region | 2017-2020",
       subtitle = "Number of people (million)",
       x = "",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_y_continuous(expand = expansion(c(0, 0.2)),
                     breaks = pretty_breaks()) +
  theme_unhcr(grid = FALSE,
              axis = "X") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank())
```

## Grouped bar chart


```{r col2, fig.width = 5.5, fig.height = 5}
df4 <- tribble(
  ~region,~'2019',~'2020',
  "West and Central Africa",70834,21581,
  "Southern Africa",85960,18022,
  "Middle East/North Africa",70799,35805,
  "Europe",914699,649308,
  "East and Horn of Africa,\nand Great Lakes",71303,47924,
  "Asia and Pacific",119894,58789,
  "Americas",976547,480286)

df4 |>
  pivot_longer(cols = -region) |>
  ggplot() +
  geom_col(aes(value/1e3, fct_rev(factor(region)), fill = name),
           position = position_dodge(0.7),
           width = 0.6) +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  labs(title = "Individual asylum applications registered by region | 2019-2020",
       x = "Number of people (thousand)",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_fill_unhcr_d(palette = "pal_unhcr") +
  theme_unhcr(font_size = 9,
              grid = "X",
              axis = "y") +
  theme(axis.title.y = element_blank())
```

## Grouped bar chart with data label

```{r col2_alt, fig.width = 5.5, fig.height = 5}
df4 |>
  pivot_longer(cols = -region) |>
  ggplot() +
  geom_col(aes(value/1e3, fct_rev(factor(region)), fill = name),
           position = position_dodge(width = 0.7),
           width = 0.6) +
  geom_text(aes(value/1e3,
                fct_rev(factor(region)),
                group = name,
                label = round(value/1e3)),
            position = position_dodge(width = 0.7),
            hjust = -0.25,
            size = 3) +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  labs(title = "Individual asylum applications registered by region | 2019-2020",
       subtitle = "Number of people (thousand)",
       x = "",
       y = "",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_fill_unhcr_d(palette = "pal_unhcr") +
  theme_unhcr(font_size = 9,
              grid = FALSE,
              axis = "y") +
  theme(axis.title.y = element_blank())
```

## Stacked column chart

```{r bar3, fig.width = 5.5, fig.height = 3.5}
df5 <- tribble(
~Year,~`UNHCR resettlement`,~`Other resettlement`,
"2010",73,26,
"2011",62,18,
"2012",69,20,
"2013",71,27,
"2014",74,32,
"2015",82,26,
"2016",126,47,
"2017",65,38,
"2018",56,37,
"2019",64,44,
"2020",23,11)

df5 |>
  pivot_longer(cols = -Year) |>
  ggplot() +
  geom_col(aes(x = Year, y = value, fill = name),
           width = 0.7) +
  scale_fill_unhcr_d(palette = "pal_unhcr", nmax = 2, order = 2:1) +
  scale_y_continuous(expand = expansion(c(0, 0.25))) +
  labs(title = "Resettlement by UNHCR and others | 2010-2020",
       x = "",
       y = "",
       caption = "Source: data source here\n© UNHCR, The UN Refugee Agency") +
  theme_unhcr(font_size = 9,
              grid = "Y",
              axis = "X")
```

## Stacked column chart with data label


```{r bar3_alt, fig.width = 5.5, fig.height = 3.5}
df5 |>
  pivot_longer(cols = -Year) |>
  ggplot() +
  geom_col(aes(x = Year, y = value, fill = name),
           width = 0.7,
           position = position_stack()) +
  geom_text(aes(x = Year, y = value, color = name, label = value),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE,
            size = 3) +
  scale_color_manual(values = c("#000000", "#FFFFFF")) +
  scale_fill_unhcr_d(palette = "pal_unhcr", nmax = 2, order = 2:1) +
  scale_y_continuous(expand = expansion(c(0, 0.25))) +
  labs(title = "Resettlement by UNHCR and others | 2010-2020",
       subtitle = "Number of people (thousand)",
       x = "",
       y = "",
       caption = "Source: data source here\n© UNHCR, The UN Refugee Agency") +
  theme_unhcr(font_size = 9,
              grid = FALSE,
              axis = "X") +
  theme(axis.text.y = element_blank())
```

## Stacked bar chart

```{r bar4, fig.width = 5.5, fig.height = 5}
df6 <- tribble(
~Country,~Refugees,~`Asylum-seekers`,~`Venezuelans displaced abroad`,
"Eritrea",0.5,0.1,0.0,
"Central African Rep.",0.6,0.0,0.0,
"Sudan",0.8,0.1,0.0,
"Somalia",0.8,0.1,0.0,
"DRC",0.8,0.1,0.0,
"Myanmar",1.1,0.0,0.0,
"South Sudan",2.2,0.0,0.0,
"Afghanistan",2.6,0.2,0.0,
"Venezuela",0.2,0.9,3.8,
"Syria",6.7,0.1,0.0)

df6 |>
  pivot_longer(cols = -Country) |>
  ggplot() +
  geom_col(aes(x = value, y = reorder(Country, value), fill = name),
           width = 0.7) +
  labs(title = "People displaced across borders by country of origin | 2020",
       x = "Number of people (million)",
       y = "",
       caption = "Source: data source here\n© UNHCR, The UN Refugee Agency") +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  scale_fill_unhcr_d(palette = "pal_unhcr",
                     nmax = 10, order = c(3, 1, 9)) +
  theme_unhcr(font_size = 9,
              grid = "X",
              axis = "y")
```

## Stacked bar chart with data label

```{r bar4_alt, fig.width = 5.5, fig.height = 5}
df6_long <- df6 |>
  pivot_longer(cols = -Country)

ggplot() +
  geom_col(data = df6_long,
           aes(x = value, y = reorder(Country, value), fill = name),
           width = 0.7,
           position = position_stack()) +
  geom_text(data = filter(df6_long, value > 0),
            aes(x = value, y = reorder(Country, value),
                group = name, label = value),
            position = position_stack(vjust = 0.5),
            color = "#FFFFFF",
            size = 3) +
  labs(title = "People displaced across borders by country of origin | 2020",
       subtitle = "Number of people (million)",
       x = "",
       y = "",
       caption = "Source: data source here\n© UNHCR, The UN Refugee Agency") +
  scale_x_continuous(expand = expansion(c(0, 0.1))) +
  scale_fill_unhcr_d(palette = "pal_unhcr",
                     nmax = 10, order = c(3, 1, 9)) +
  theme_unhcr(font_size = 9,
              grid = FALSE,
              axis = "y") +
  theme(axis.text.x = element_blank())
```

## 100% stacked column chart

```{r bar5, fig.width = 5.5, fig.height = 3}
df7 <- tribble(
~Year,~Unearmarked,~`Softly earmarked`,~Earmarked,~`Tightly earmarked`,
"2012",20,32,20,28,
"2013",16,20,30,34,
"2014",15,18,37,30,
"2015",14,22,40,24,
"2016",15,20,43,22,
"2017",15,20,45,20,
"2018",15,18,40,27,
"2019",16,14,45,25,
"2020",14,15,46,25)

df7_long <- df7 |>
  pivot_longer(cols = -Year)

df7 |>
  pivot_longer(cols = -Year) |>
  ggplot() +
  geom_col(aes(x = Year, y = value, fill = name),
           width = 0.7,
           position = position_fill()) +
  scale_fill_unhcr_d(palette = "pal_unhcr") +
  scale_y_continuous(expand = expansion(c(0, 0.01)),
                     labels = percent) +
  labs(title = "Levels of earmarking | 2012-2020",
       x = "",
       y = "",
       caption = "Source: UNHCR\n© UNHCR, The UN Refugee Agency") +
  theme_unhcr(font_size = 9,
              grid = "Y",
              axis = "X")
```

## Population pyramid

```{r bar6, fig.width = 5.5, fig.height = 5}
df8 <- tribble(
  ~ages,~male,~female,
  "0-4",0.06,0.06,
  "5-11",0.09,0.08,
  "12-17",0.06,0.06,
  "18-59",0.27,0.25,
  "60+",0.03,0.03) |>
  mutate(ages = factor(ages,
                       levels = c("0-4", "5-11", "12-17", "18-59", "60+"),
                       ordered = TRUE))


ggplot() +
  geom_col(data = df8,
           aes(-male, ages, fill = "Male"),
           width = 0.7) +
  geom_col(data = df8,
           aes(female, ages, fill = "Female"),
           width = 0.7) +
  geom_text(data = df8,
            aes(-male, ages, label = percent(abs(male))),
            hjust = 1.25,
            size = 3) +
  geom_text(data = df8,
            aes(female, ages, label = percent(female)),
            hjust = -0.25,
            size = 3) +
  labs(title = "Demographics of people who have been forcibly displaced | 2020",
       caption = "Note: figures do not add up to 100 per cent due to rounding
Source: UNHCR Refugee Data Finder
© UNHCR, The UN Refugee Agency") +
  scale_x_continuous(expand = expansion(c(0.1, 0.1))) +
  scale_fill_manual(values = setNames(unhcr_pal(n = 3, "pal_unhcr")[c(2, 1)],
                                      c("Male", "Female"))) +
  theme_unhcr(font_size = 9,
              grid = FALSE,
              axis = FALSE) +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

## Line chart

```{r line, fig.width = 5.5, fig.height = 3}
df9 <- tribble(
~Year,~Refugees,~IDPs,
1990,17.4,0,
1991,16.85,0,
1992,17.84,0,
1993,16.33,4.2,
1994,15.75,5.32,
1995,14.9,4.29,
1996,13.36,4.85,
1997,12.01,4.57,
1998,11.48,5.06,
1999,11.69,3.97,
2000,12.13,6,
2001,12.12,5.1,
2002,10.59,4.65,
2003,9.59,4.18,
2004,9.57,5.43,
2005,8.66,6.62,
2006,9.88,12.79,
2007,11.39,13.74,
2008,10.49,14.44,
2009,10.4,15.63,
2010,10.55,16,
2011,10.4,15.47,
2012,10.5,17.67,
2013,11.7,23.93,
2014,14.38,32.27,
2015,16.11,37.49,
2016,17.18,36.63,
2017,19.94,39.12,
2018,20.36,41.43,
2019,20.41,43.5,
2020,20.65,48.56)

df9 |>
  pivot_longer(cols = c("Refugees", "IDPs")) |>
  ggplot(aes(x = Year, y = value, color = name)) +
  geom_line(size = 1) +
  geom_dl(aes(label = name),
          method = list(dl.trans(x = x + 0.1),"last.points"),
          size = 3) +
  labs(title = "Number of refugees and IDPs of concern to UNHCR | 1990-2020",
       y = "Number of people (million)",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_y_continuous(expand = expansion(c(0, 0.1))) +
  scale_x_continuous(expand = expansion(c(0, 0.25))) +
  scale_color_unhcr_d(palette = "pal_unhcr",
                      nmax = 10, order = c(5, 1)) +
  coord_cartesian(clip = "off") +
  theme_unhcr(font_size = 9,
              grid = "Y") +
  theme(legend.position = "none",
        axis.title.y = element_blank())
```

## Area chart

```{r area, fig.width = 5.5, fig.height = 3}
df10 <- tribble(
~Month,~Unearmarked,~`Softly earmarked`,~Earmarked,~`Tightly earmarked`,
"Jan",15.1,0.6,1.6,19.2,
"Feb",15.3,3.9,1.7,21.9,
"Mar",11.9,7.6,20.1,22.8,
"Apr",15.7,9.8,41.8,39.8,
"May",22.8,14.8,34,51.3,
"Jun",29.1,18,34.6,53.7,
"Jul",41.5,61.9,126.6,63.7,
"Aug",58,61.6,118.1,92.9,
"Sep",62.6,78.2,130.1,86,
"Oct",69.3,81.3,146.6,88.8,
"Nov",76.6,83.5,145.7,95.8,
"Dec",99.8,102.1,145.7,107.3)

df10 |>
  pivot_longer(cols = -Month) |>
  mutate(ym = paste0("01-", Month, "-2021"),
         ym = dmy(ym)) |>
  ggplot() +
  geom_area(aes(ym, value, fill = name)) +
  scale_fill_unhcr_d(palette = "pal_unhcr") +
  labs(title = "Monthly evolution of funding in West and Central Africa region | 2020",
       caption = "Source: UNHCR Refugee Data Finder\n© UNHCR, The UN Refugee Agency") +
  scale_x_date(breaks = seq.Date(as.Date("2021-01-01"),
                                 as.Date("2021-12-01"),
                                 by = "1 month"),
               date_labels = "%b",
               expand = expansion(c(0, 0))) +
  scale_y_continuous(expand = expansion(c(0, 0.1))) +
  theme_unhcr(font_size = 9,
              grid = "Y")
```

## Combined column and line chart

```{r line2, fig.width = 5.5, fig.height = 3}
df11 <- tribble(
~Year,~`Displaced population`,~`Proportion displaced`,
2007,42.69,6.38964,
2008,41.98,6.20603,
2009,43.25,6.31717,
2010,43.71,6.30724,
2011,42.5,6.0596,
2012,45.16,6.36237,
2013,51.2,7.12979,
2014,59.53,8.19344,
2015,65.3,8.88499,
2016,65.36,8.82591)


ggplot(data = df11) +
  geom_col(aes(Year,
               `Displaced population`, fill = "Displaced population"),
           width = 0.7) +
  geom_line(aes(Year,
                `Proportion displaced` * 10, color = "Proportion displaced"),
            size = 1) +
  scale_fill_manual(values = setNames(unhcr_pal(n = 1, "pal_blue"),
                                      "Displaced population")) +
  scale_color_manual(values = setNames(unhcr_pal(n = 1, "pal_red"),
                                       "Proportion displaced")) +
  labs(title = "Trend of global displacement | 2007 - 2016",
       caption = "Source: UNHCR Refugee Data Finder
© UNHCR, The UN Refugee Agency") +
  scale_x_continuous(breaks = pretty_breaks(n = nrow(df11))) +
  scale_y_continuous("Displaced population (millions)",
                     expand = expansion(c(0, 0.2)),
                     breaks = pretty_breaks(),
                     sec.axis = sec_axis(~ . / 10,
                                         breaks = pretty_breaks(),
                                         name = "Proportion displaced (number displaced per 1,000)")) +
  theme_unhcr(font_size = 9,
              grid = "Y") +
  theme(axis.title.x = element_blank())
```

## Treemap

```{r treemap, fig.width = 5.5, fig.height = 5}
df13 <- tribble(
  ~region,~workforce,
  "West and Central Africa",5,
  "Southern Africa",11,
  "Middle East/North Africa",19,
  "Europe",10,
  "East and Horn of \nAfrica, and Great Lakes",22,
  "Asia and Pacific",12,
  "Americas",11,
  "HQ", 9)

df13 |>
  mutate(label = paste0(workforce, "%\n", region)) |>
  ggplot(aes(area = workforce)) +
  geom_treemap(color = "#FFFFFF",
               size = 1,
               fill = unhcr_pal(n = 1, "pal_blue"),
               start = "topleft") +
  geom_treemap_text(aes(label = label),
                    color = "#FFFFFF",
                    grow = FALSE,
                    start = "topleft",
                    family = "Lato") +
  labs(title = "UNHCR global workforce by region | 2020",
       caption = "Note: figures do not add up to 100 per cent du to rounding
Source: UNHCR Refugee Data Finder
© UNHCR, The UN Refugee Agency") +
  theme_unhcr(font_size = 9,
              axis = FALSE,
              grid = FALSE)
```


## Pie chart

```{r pie, fig.width = 5.5, fig.height = 5}
df12 <- tribble(
  ~label,~Funding,
  "Received to date",144,
  "Gap",150)

df12 |>
  mutate(p = Funding / sum(Funding),
         label2 = paste0("<strong>$", Funding, " M</strong>",
                         "<br>", label, " (",
                         round(100 * p, 1), "%)")) |>
  ggplot() +
  geom_arc_bar(aes(x0 = 0,
                   y0 = 0,
                   r0 = 0,
                   r = 1,
                   amount = Funding,
                   fill = label),
               size = 1,
               color = "#FFFFFF",
               stat = "pie") +
  geom_richtext(x = c(1.45, -1.25), y = c(-0.3, 0.3),
                aes(label = label2),
                size = 3,
                fill = NA,
                label.color = NA) +
  scale_fill_unhcr_d(palette = "pal_unhcr") +
  scale_x_continuous(expand = expansion(c(0.3, 0.5))) +
  coord_fixed(clip = "off") +
  theme_unhcr(grid = FALSE, axis = FALSE) +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank())
```

## Donut chart

```{r donut, fig.width = 5.5, fig.height = 5}
df12 |>
  mutate(p = Funding / sum(Funding),
         label2 = paste0("**$", Funding, " M**",
                         "<br>", label, " (",
                         round(100 * p, 1), "%)")) |>
  ggplot() +
  geom_arc_bar(aes(x0 = 0,
                   y0 = 0,
                   r0 = 0.65,
                   r = 1,
                   amount = Funding,
                   fill = label),
               size = 1,
               color = "#FFFFFF",
               stat = "pie") +
  geom_richtext(x = c(1.45, -1.25), y = c(-0.3, 0.3),
                aes(label = label2),
                size = 3,
                fill = NA,
                label.color = NA) +
  geom_richtext(x = 0, y = 0,
                label = paste0("**$", sum(df12$Funding), " M**",
                               "<br><span style='font-size:12pt'>Total required</span>"),
                fill = NA,
                label.color = NA,
                size = 10) +
  scale_fill_unhcr_d(palette = "pal_unhcr") +
  scale_x_continuous(expand = expansion(c(0.3, 0.5))) +
  coord_fixed(clip = "off") +
  theme_unhcr(grid = FALSE, axis = FALSE) +
  theme(legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank())
```
