---
title: "Designing UNHCR Branded Charts using unhcrthemes"
---

```{r, include = FALSE}
library(ragg)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  fig.retina = 2,
  dpi = 300
)
```

Data visualization is a powerful tool for conveying information and insights from complex datasets. The [United Nations High Commissioner for Refugees](https://www.unhcr.org/) (UNHCR) provides crucial [refugee statistics](https://www.unhcr.org/refugee-statistics/download/) that can be presented in an informative and visually appealing manner. The `{unhcrthemes}` package, in combination with the `{refugees}` package and the versatile `{ggplot2}` library, offers an excellent solution to create UNHCR-branded charts with ease.

## Setup

Before we start creating UNHCR branded charts, make sure you have the **necessary packages installed**:

```{r install, eval=FALSE}
# Install CRAN packages
packages <- c("refugees", "tidyverse", "scales", "ragg", "pak")
install.packages(setdiff(packages, rownames(installed.packages())))

# Install unhcrthemes with pak
pak::pkg_install("unhcr-dataviz/unhcrthemes")
```

**Load the packages** into your R environment:

```{r setup}
# Load required packages
library(unhcrthemes)
library(refugees)
library(tidyverse)
```

Some packages installed above will be loaded using the `::` notation to make the code and intentions clearer.

For this tutorial, we'll use statistics provided by UNHCR. The refugee package allows us to access and work with this data. You can find more information on the datasets available and the use of the `{refugees}` package in [this blog post](https://www.unhcr.org/refugee-statistics/insights/explainers/refugees-r-package.html).

## Bar chart

Let's start by creating a simple bar chart to visualize the top 10 refugee and other people in need of international protection by country of origin in 2022.

**Preparing the data:**

```{r bar-data}
# Keep only 2022 records, Sum refugees and oip,
# Keep only top 10, Wrap long label
bar_df <- refugees::population |>
  filter(year == 2022) |>
  summarise(
    refugees = sum(refugees, na.rm = TRUE) + sum(oip, na.rm = TRUE),
    .by = coo_name
  ) |>
  slice_max(order_by = refugees, n = 10) |>
  mutate(coo_name = str_wrap(coo_name, 25))
```

**Plot:**

```{r bar-plot, out.width="100%"}
# Set title, subtitle and caption
# Caption will be used throughout the tutorial
title_bar <- "Refugees and other people in need of international protection"
subtitle_bar <- "By country of origin at the end of 2022"
caption <- "Source: Refugee Data Finder<br>© UNHCR, The UN Refugee Agency"

# Plot
ggplot(
  bar_df,
  aes(
    x = refugees,
    y = reorder(coo_name, refugees)
  )
) +
  geom_col(
    fill = unhcr_pal(n = 1, "pal_blue"),
    width = 0.8
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, .1)),
    breaks = seq(0, 7e6, by = 1e6),
    labels = scales::label_number(scale_cut = scales::cut_short_scale())
  ) +
  labs(
    title = title_bar,
    subtitle = subtitle_bar,
    caption = caption,
    x = "Number of people"
  ) +
  theme_unhcr(grid = "X", axis = "Y", axis_title = "X")
```

**Plot with labels:**

```{r bar-label, out.width="100%"}
# Plot with labels
ggplot(
  bar_df,
  aes(
    x = refugees,
    y = reorder(coo_name, refugees)
  )
) +
  geom_col(
    fill = unhcr_pal(n = 1, "pal_blue"),
    width = 0.8
  ) +
  geom_text(
    aes(
      label = scales::label_number(
        scale_cut = scales::cut_short_scale(),
        accuracy = .1
      )(refugees)
    ),
    hjust = -.2
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(
    title = title_bar,
    subtitle = subtitle_bar,
    caption = caption
  ) +
  theme_unhcr(grid = FALSE, axis = "Y", axis_text = "Y", axis_title = FALSE)
```

## Column chart

A column chart is quite similar to a bar chart, but the bars are vertical. Let's make a column chart of the last 10 years of total refugee population.

**Preparing the data:**

```{r column-data}
# Keep only 2013-2022 records, Sum refugees, Make year factor
column_df <- refugees::population |>
  filter(year >= 2013) |>
  summarise(
    refugees = sum(refugees, na.rm = TRUE),
    .by = year
  ) |>
  mutate(year = as_factor(year))
```

**Plot:**

```{r column-plot, out.width="100%"}
# Set title
# Caption come from the bar chart example
title_column <- "Total number of refugees | 2013-2022"

# Plot
ggplot(
  column_df,
  aes(x = year, y = refugees)
) +
  geom_col(
    fill = unhcr_pal(n = 1, "pal_blue"),
    width = 0.8
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, .1)),
    breaks = seq(0, 30e6, by = 5e6),
    labels = scales::label_number(scale_cut = scales::cut_short_scale())
  ) +
  labs(
    title = title_column,
    caption = caption,
    Y = "Number of people"
  ) +
  theme_unhcr(grid = "Y", axis = "X", axis_title = FALSE)
```

**Plot with labels:**

```{r column-label, out.width="100%"}
# Plot with labels
ggplot(
  column_df,
  aes(x = year, y = refugees)
) +
  geom_col(
    fill = unhcr_pal(n = 1, "pal_blue"),
    width = 0.8
  ) +
  geom_text(
    aes(
      label = scales::label_number(
        scale_cut = scales::cut_short_scale(),
        accuracy = .1
      )(refugees)
    ),
    vjust = -.4
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, .1)),
    breaks = seq(0, 30e6, by = 5e6),
    labels = scales::label_number(scale_cut = scales::cut_short_scale())
  ) +
  labs(
    title = title_column,
    caption = caption,
    Y = "Number of people"
  ) +
  theme_unhcr(grid = FALSE, axis = "X", axis_title = FALSE, axis_text = "X")
```
